Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-07 p2/ordering.tex
New definitions based on equivalence classes

@@

\begin{mydef}[$\ordersetsign$]
We use $\orderset{U}$ to denote the set of sequences that contain all commands in
the simple set $U$
and represent all orders generated by this algorithm.
\end{mydef}

It follows from the algorithm that:
\begin{mycor}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a $U'\in\orderset{U}$ where they are next to each other.
\end{mycor}

As a specific case, we can regard a simple sequence of commands $S$
as a set and determine its $\orderset{S}$.
\Cref{simple_reorder_equiv} is
an important property of simple sequences: namely, that
all of their valid orderings are equivalent,
and all equivalent sequences are mere reorderings.
To prove this, first we prove the following:

@@

\begin{mylem}\label{simple_reorder_equiv}
For all simple sequences $T$ and $S\nequiv\cbrk$,
\[ T\equiv S \Longleftrightarrow T\in\orderset{S}. \]
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

@@

>>>>

\begin{mydef}[$\ordersetsign$]
We use $\orderset{U}$ to denote the set of sequences 
generated by the algorithm / (all possible orderings)
from the commands in the simple-? set or sequence $U$.
\end{mydef}

@@

\noindent As a specific case, we can regard a simple sequence $S$
as a set, and use the above algorithm to generate all possible orderings
of the commands it contains.
\Cref{simple_reorder_equiv} is
an important property of simple sequences: namely, that
all of their valid reorderings are equivalent,
and all equivalent sequences are mere reorderings.
To prove this, first we prove the following:

@@

\begin{mylem}\label{simple_reorder_equiv}
For a simple sequence $S\nequiv\cbrk$,
\[ \orderset{S} = [S]_\equiv, \]
where $[S]_\equiv$ denotes the set of simple sequences equivalent to $S$.
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

@@

%% TODO https://en.wikipedia.org/wiki/Equivalence_class

It follows from the algorithm that:
\begin{mycor}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a simple $U'\equiv U$ in which they are next to each other.
\end{mycor}

### 2016-10-07 p2/works-inverse.tex
New definitions based on equivalence classes

\noindent
We continue by defining inverse commands and sequences
which allow us to move parts of sequences between the
condition and consequence parts of $\workssign$.

\begin{mydef}[Inverse commands and sequences]
The inverse of command $\cxynv$ is $\cxynv^{-1} = \caaaa{Y}{X}{n}{\valvx}$
where $\valvx$ is an arbitrary value from $\setvx{X}$.
We write $S^{-1}$ for the inverse of sequence $S$, which consists of the inverses of the commands in $S$
in reverse order.
\end{mydef}

\begin{mydef}[$\typeeq$]
For two filesystems we write $\FS\typeeq\GS$
iff 
at all nodes the types of their values are the same, or they are both broken.
\end{mydef}

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq}
$\forall S, \FS: (S^{-1})^{-1}\aFS\typeeq S\aFS$.
\end{mycor}

\begin{mylem}\label{r_invmove}
A common initial segment of a set of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \\
\worksc{\seqset{A}}{B\cc \seqset{C}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{C}}
\end{gather*}
\end{mylem}
\begin{proof}
This is based on the fact that in our model, unless they break a filesystem,
commands leave the type of the filesystem value unchanged 
or change one type into another, but never merge types.
That is,
sequences---as functions mapping filesystems to filesystems---are
essentially bijections over type-equality ($\typeeq$)
with the only ``sink'' being $\fsbroken$:
\begin{align*}
\forall S,\FS,\GS:{}&S\aFS\neq\fsbroken \Rightarrow \\
&(S\aFS\ntypeeq S\aGS \Longleftrightarrow \FS\ntypeeq \GS).
\end{align*}

>>>>

%% TODO https://en.wikipedia.org/wiki/Equivalence_class

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems we write $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes,
and at all nodes the types of their values are the same.
\item For two commands we write $\alpha\typeeq\beta$ iff their input and output types,
and the node they operate on are the same.
For example, $\cbfaa{n}{f_1}\typeeq\cbfaa{n}{f_2}$.
\item For two sequences of commands we write $A\typeeq B$ iff they contain the same number of commands,
and pairwise the commands $\alpha\in A$ and $\beta\in B$ at the same indices are equivalent by type ($\alpha\typeeq\beta$).
\end{itemize}
\end{mydef}

In other words, two type-equivalent objects can only differ in the value they contain,
but as $|\setb|=|\setd|=1$, this is only possible at values of the type $\ccharf$.
We select an arbitrary file value $\fil\in\setf$, and choose a representative
of each $\typeeq$-equivalence class where all (if any) values of the type $\ccharf$ is $\fil$.
For example, in the case of commands, the representative of $[\cbfaa{n}{f}]_\typeeq$ is $\cbfaa{n}{\fil}$.

\begin{mydef}[$\T$]
We use $\T$ and write $x\T$ to denote the canonical mapping of 
filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
$\T$ is also extended to sets of sequences,
where $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{mydef}

For example, $\cbfaa{n}{f}\T = \cbfaa{n}{\fil}$,
and for the sequence $S=\alpha\cc\beta$, $S\T=\alpha\T\cc\beta\T$.
In other words, $\T$ maps our system of filesystems and commands
into a system where $|\setf|$ is also $1$.

\begin{mycor}
For any filesystem $\FS$ and command $\alpha$, 
$\alpha\T(\FS\T) = (\alpha\aFS)\T$.
\end{mycor}

% It is easy to see that 
% $A\equiv B \Rightarrow A\T\equiv B\T$ 
% and
% $A\eqext B \Rightarrow A\T\eqext B\T$,
% but the reverse is not true as 
% $A\T$ and $B\T$ will have the same effect on all filesystems
% if $A$ and $B$ differ only in some of the values in their commands,
% whereas clearly $A\nequiv B$.

As whether a sequence of commands breaks a filesystem does not depend on the output values
of the commands, only the input and output types, we know that
\begin{mycor}
\begin{gather*}
\works{\seqset{A}} \Leftrightarrow \works{\seqset{A}\T} \\
\worksc{\seqset{A}}{\seqset{B}} \Leftrightarrow \worksc{\seqset{A}\T}{\seqset{B}\T}.
\end{gather*}
\end{mycor}

We are now ready to define the inverse of representative commands $\alpha\T$
which allows us to move parts of sequences between the
condition and consequence parts of $\workssign$.
This is based on the following observation.

\begin{mylem}
Representative functions $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying the command to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
We know that a representative filesystem can only have values from the set $\rvalueset$.
As $\alpha\T(\FS\T)\neq\fsbroken$, $\FS\T(n)$ must be the value that matches the input type of $\alpha\T$.
We know that $\GS\T(n)\neq\FS\T(n)$, and therefore $\GS\T(n)$ must have a different type,
which implies $\alpha\T(\GS\T)=\fsbroken$ as it cannot match the input type of $\alpha\T$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

\begin{mydef}[Inverse commands and sequences]
If we regard the representative command $\alpha\T$ a function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq} %% TODO Don't use!
$\forall S, \FS: (S^{-1})^{-1}\aFS\typeeq S\aFS$.
\end{mycor}

\begin{mylem}\label{r_invmove}
A common initial segment of a set of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \\
\worksc{\seqset{A}}{B\cc \seqset{C}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{C}}
\end{gather*}
\end{mylem}
\begin{proof}

%% TODO see lemma above

This is based on the fact that in our model, unless they break a filesystem,
commands leave the type of the filesystem value unchanged 
or change one type into another, but never merge types.
That is,
sequences---as functions mapping filesystems to filesystems---are
essentially bijections over type-equality ($\typeeq$)
with the only ``sink'' being $\fsbroken$:
\begin{align*}
\forall S,\FS,\GS:{}&S\aFS\neq\fsbroken \Rightarrow \\
&(S\aFS\ntypeeq S\aGS \Longleftrightarrow \FS\ntypeeq \GS).
\end{align*}
